### 2019-04-17 Plan/plan.tex

\documentclass[a4paper,smallheadings,11pt,oneside,bibliography=totoc]{scrreprt}

%Used to change the margin
\newcommand{\margin}{0.75in}
\usepackage[a4paper, left=\margin, right=\margin, top=\margin, bottom=\margin]{geometry}

% Packages
\usepackage[numbers]{natbib} 		% Package for IEE Referencing
\usepackage{graphicx} 				% Package for images
\usepackage{hyperref} 				% Package for URLS and Content links
\usepackage[titletoc]{appendix} 	% Package added for the appendix
\usepackage{mathtools}				% Package for mathmatical notation
\usepackage{wrapfig}				% Wrapping figures on a page
\usepackage{tcolorbox}				% For code listings
\usepackage{titlesec}				% Package for reducing space after headers

%Paragraph settings
\setlength{\parindent}{0em}
\setlength{\parskip}{0.75em}

\begin{document}

	\input{pages/work_plan}
	\input{pages/delivery_plan}

\end{document}