### 2019-05-19 Papers/Notes_TeX/pages/paper_notes.tex
Changes; - Restructuring of the archive folder - Added TeX version of paper notes - Added new papers for relevancy review - Included a script to generate fabricated target_key lists for benchmarking

\textbf{TODO:} Need to find papers on study of words are a comparative measure.

\section{Can Unicorns Help Users Compare Crypto Key Fingerprints}

This paper tested 8 different fingerprint representations with 661 different participants. All of these representations are testing using compare-and-confirm (basic comparison process for fingerprints) and compare-and-select (select from a list)

\begin{itemize}
    \item Hexadecimal
    \item TODO
\end{itemize}

The paper also emphasises the attention that has been taken to realism in the experimentation stage.

Talks about long term usability issues with public-key crypto and states fingerprints are a main source of the problem. ([28])

There seems to be mixed findings with compare-and-select and compare-and-confirm ([14, 17])

\subsection{Findings}

\begin{itemize}
    \item Graphical representations have mixed success rates (with quick comparison times)
    \item Recommendations not to use \textbf{compare-and-select}
\end{itemize}