### 2019-05-27 Papers/Notes_TeX/pages/improving_robustness.tex

\section{Improving the Robustness of Wireless Device
Pairing Using Hyphen-Delimited Numeric
Comparison}

In the context of secure device pairing the paper aims to improve the human interaction with number pairing.

Lots of research has gone into Short Authentication String (SAS) numbers ([14], [10], [4])

The research was peformed on very old devices. Nokia 6030b, very out of date and not reflective of anything around at this time.

Usability study looked into Efficiency, Robustness, General Usability

Strange format of keeping the numbers exactly the same but just placing hyphens in the middle. The authors attempted to remedy this by spacing the experiments with a days gap

Study was performed with 40 participants

Very small number of values tested only 5 overall???

Initial results show promising results for hypen delimitations etc alongside improved user impressions