### 2019-05-29 Documents/paper_notes/pages/use_your_words.tex

\section{Use Your Words: Designing One-time Pairing Codes toImprove User Experience}

The use of works instead of numerical codes. This is in the context of device pairing in IoT. Something that is new in the literature

Search space required was 500,000,000

Numeric:        9 digits
Alphanumeric:   5 chars
Words:          3 words

Paper contains references to lots of research into words distributions etc

Paper claimed that 800 works would need to be used but in the experiment only utilised 242 words???

Words provided a faster entry time with no reduction in accuracy

``It is known that typing familiar words is faster than typing random strings [22]''

Paper did not consider malicious aspect of their methods, no discussion was made into how a dictionary of 800 words could be easily impersonated

However, promising results for the usability of word based encodings on displays with low fidelity