### 2019-05-31 Documents/literature_review/tables/fingerprint_attacker_stren.tex
Changes; - Quantified accuracy with a table - Added results for speed of comparison - Small wording changes

\begin{tabular}{c|ccccc}
    \toprule
                        & Hsu-Chun\cite{hsiao2009study}      
                        & Kainda\cite{kainda2009usability}      
                        & Dechand\cite{dechand2016empirical}
                        & Tan\cite{tan2017can}      
                        & M. Shirvanian\cite{shirvanian2017pitfalls}
                        \\\hline
    Attacker Strength\footnote{Some attacker strengths are estimations due to the lack of quantification in some papers.}   
    & $\sim2^{28}$  & $\sim2^{40}$ & $2^{80}$ & $2^{60}$ & $\sim2^{242}$ \\
    Entropy Range       & 22-28 bits    & 20-40 bits   & 122 bits & 128 bits & 160-256 bits  \\
    No\degree  ~Participants     & 436           & 30           & 1047     & 661      & 25            \\
    \bottomrule
\end{tabular}