### 2019-06-03 Documents/literature_review/appendix.tex
Changes; - Added more content to the new lit' review structure

\begin{appendices}
    \section{Encoding Scheme Results}
    \begin{table}[h!]
        \makebox[\textwidth][c]{
            \input{tables/fingerprint_comparison.tex}
        }%
        \caption{Accuracy of correct comparison for the encoding schemes assessed}
        \label{tab:results}
    \end{table}

    \begin{table}[h!]
        \makebox[\textwidth][c]{
            \input{tables/fingerprint_times.tex}
        }%
        \caption{Timing results in seconds for the related schemes}
        \label{tab:times}
    \end{table}
\end{appendices}