### 2019-07-22 Thesis/pages/background.tex

\chapter{Background}
\label{cha:Background}

\section{PGP}

\section{OpenCL}

\section{Pretty Easy Privacy/Trustwords}