### 2019-07-30 Thesis/pages/introduction.tex

\chapter{Introduction}
\label{cha:Introduction}

The increasing use of public key cryptography by instant messaging and secure email means key fingerprint  verification is an ever more important task. One of the most significant risks to the security of the communication channel is a Man-in-the-middle (MiTM) attack. A successful MiTM attack can circumvent the encryption as it allows an attacker to read all encrypted data. A countermeasure for this is the verification of each parties' fingerprint. Fingerprints can come in a number

\textbf{TODO: } Listing of chapters and how the report is to be laid out