### 2019-08-02 Thesis/pages/experiments.tex
Changes; - Small changes made to the main experiment - Updated lists of vuln keys - Added more content to the thesis

\chapter{Experiments}
\label{cha:Experiments}

\section{Scallion vs GreenOnion}
% Show that the bloom filter blows Scallion out of the water

\section{Experiment 1}
% Demographic

\begin{table}[h!]
    \centering
    \begin{tabular}{|l|ll|}
        \hline
        Gender & Male: & 46.2\% \\
               & Female: & 53.8\% \\
        \hline
        Age:   & 18-24: & 10.6\% \\ 
               & 25-29: & 20.2\% \\ 
               & 30-39: & 30.8\% \\ 
               & 40-49: & 22.1\% \\ 
               & 50-59: & 11.5\% \\ 
               & 60-69: & 3.8\% \\ 
               & 70-79: & 1.0\% \\

        \hline
        Highest Education:  & Bachelor's degree:    & 51.0\% \\
                            & A-Level/O-Level:      & 18.3\% \\
                            & GCSE:                 & 15.4\%  \\
                            & Master's degree:      & 13.5\% \\ 
                            & PhD:                  & 1.9\% \\
        \hline

    \end{tabular}
    \caption{Participant demographics}
    \label{}
\end{table}

% Results

\section{Experiment 2}
% Methodology
% Demographic
% Results