### 2019-08-02 Thesis/pages/implementation.tex

\chapter{Implementation}
\label{cha:Implementation}

\textbf{TODO: } Talk here about the interesting and challenges that were overcome

\section{GreenOnion}
Talk about the way a key is expanded and turned into a lot
of work for the kernel

\section{First Experiment}
Explain how Google App Script was used?

\section{MainExperiment}
Design of the application etc