### 2019-08-06 Thesis/diagrams/graphs/experiment_one/graph.tex
Changes; - Added graphs for firstexperiment results - More content added to thesis

\begin{figure}[!h]
  \minipage{0.5\textwidth}
    \input{diagrams/graphs/experiment_one/soundex.tex}
    \caption{Soundex}
  \endminipage
  \minipage{0.5\textwidth}
    \input{diagrams/graphs/experiment_one/leven.tex}
    \caption{Levenshtein}
  \endminipage
  \\
  \minipage{0.5\textwidth}
    \input{diagrams/graphs/experiment_one/nysiis.tex}
    \caption{NYSIIS}
  \endminipage
  \minipage{0.5\textwidth}
    \input{diagrams/graphs/experiment_one/metaphone.tex}
    \caption{Metaphone}
  \endminipage
  \\
  \centering
  \minipage{0.5\textwidth}
    \input{diagrams/graphs/experiment_one/wordvec.tex}
    \caption{Phonetic vector}
  \endminipage
  \caption{Individual breakdown of results for each metric}
  \label{fig:exp1_breakdown}
\end{figure}