### 2019-08-08 Thesis/pages/appendix.tex
Changes; - Added data from final round of experiments - Updated all batch files - Added more discussion to the expeirment section of the report

\appendix

\chapter{Randomly generated uncontrolled key}
\label{appendix:uncontrolled_key}
\begin{lstlisting}
-----BEGIN PGP PUBLIC KEY BLOCK-----

mQENBF0vHaYBCACxOy7RacHIprgaUo66vGmIB2VIaSwToliuaEBxOAloD+wTB/T9
TOKgAMLWMgK2z3/tOZj2l5k9ATUr0nZ9iBgew5Ih3ykOp+OkE9dh4NTD3EJz5UWo
M1r6scPhby92zMBqKi0iPFlPcvk+eg+SPusxIp+Vn16ALoB2pwGQvG+qoFEJfdVP
nwSAWJo1mcd+TCxgIMqe11FXDzxpd4BEpMXsHOB8NH/TUiI57sxqT8A2HnnyWo2i
5vs93VAP8rTTIRfGelW2c3oqhV9XhXbvVpJPmvkTxM/SshV1L5HhYr1TUAkAQw6w
RUZqooysgZEDqQw58lFQ7C9plCHOpXIEhbE7ABEBAAG0HEpvaG4gSm9obnNvbiA8
am9obkB3b3JrLmNvbT6JAVQEEwEIAD4WIQRQZ0ukmi5QUEz7KB3FwHKSVKqpwgUC
XS8dpgIbAwUJA8JnAAULCQgHAgYVCgkICwIEFgIDAQIeAQIXgAAKCRDFwHKSVKqp
wg4YB/9knveazSM4yN3EJY7HRZe5PVA5r0GGDlEDMte+surPKLsV/T39SNl8qTH+
DXnOyC4rt1kykTJP+ULcKqGkVRnH2XYZP/W6E9wwBPjoKHaWyn8MP0ix5kgkq6Ld
ZWEQ2pUwCsAhWv8K/ybjQ+Li484JKbrz/GRALp8qU3hZgCo/Fw3sD2KcTPU1v8qU
av9G2N5su10W7G0F1ozszFit+r2iinoZtGSddBPG+pIUEtZoBn4V2FRRSO1UY/TU
jGf2BLsk+ZkNJdx5x6IEU9CQ+ivPuKtupVu9ySKqE2Qfmn5lh9iAjYDN3XxUQfCf
braGoC9JLnL5QmT1JaQXfxkIHilyuQENBF0vHaYBCADDp/u8D/TG5d5HmabmgbBF
3R2d3J00vt7VQQ5sIRMLyP0cx6qy2bTIy9mgfU3/WcrhRp728gBo+NnUfxyiTb5R
u7AbntlkerF3dh0JLm5eIossPW37QXn3Ce4Tv4ixGRLbsGKTu6h0lP41FM5VcqV9
9NE1iml3WyngnsFdKeUYqPW4QY+ydyjpSyblbQTMptIKG2llCmsd9k+hb8nnGx/d
GvyGKaZ270iVGY2im/9VP2VwzqXlFxxx6Dux4egJJMMAy5xkzBKf+Szmm1pMJkt2
4sBrwmLnuUObgsAkq8K+MqEpWeFi6aWsh7RnY8T0SEb/3QUwG12V77ePhrarBwmf
ABEBAAGJATwEGAEIACYWIQRQZ0ukmi5QUEz7KB3FwHKSVKqpwgUCXS8dpgIbDAUJ
A8JnAAAKCRDFwHKSVKqpwoW3CACeKIcQFuVv2vpUHq/4ATH0atPYqR1NbJTegE+v
XZyrNHtmRF4V+kdx/1KX9jsIY2TjyxbhTsrtMFkMZMfAVjXVoS4y8OkwX5IJWgHC
0JQIHQrvaIGfI1yuVrNvWYyPXFxNtbntrgTd8JKYGKYaTqXNZ2r1ozbZElp7fQtR
rWNUGJYYo+aHHLM+Aot5k+Z3YqHifuPSnCEYt/GxnMhYq1IL3LSZ96hsgwqeFoLP
+v4RsvY52Yb5tiOzxZwikALOJdM2aAumxq8R1nYHiswWxotw2WkKFlB7gfLpbMWs
Y+o4J7QKSqpVTrUYBTPF1F4kBWOMY10kpeCDbZ1YwJOZJf8P
=aHBu
-----END PGP PUBLIC KEY BLOCK-----
\end{lstlisting}

\chapter{Computed Attack Keys}

\section{NYSIIS - 0 Static}
\label{appendix:nysiis_0_static}

\subsection{Controlled Key}
\begin{lstlisting}
-----BEGIN PGP PUBLIC KEY BLOCK-----

mI0EXS8lIAEEAMWEJ+NGB/ukiUijzdT+DvzVwouWV/j4Rw78X9YdgJP0ExT4ncHG
TJbOKjEMEI9bb/oYMLF/434ufScYNci+rL/3RVY8CcjEv5l0UGX+udZ78c7Z/vf1
zFr2MQAglr9M8599JwUY25LiNKzDYc87B16hqRV8hRaXHfAFFHBxRsgBABEBAAG0
HUF1dG9nZW5lcmF0ZWQgS2V5IDx1c2VyQENWRUQ+iNQEEwEKAD4WIQSlBU278MbT
iVM5h6GNohojnp5YQgUCXS8lIAIbLwUJAeJfkAULCQgHAgYVCgkICwIEFgIDAQIe
AQIXgAAKCRCNohojnp5YQvxSA/0bjL+4me3hLjrM+8IckKABCRdRYYRG7svbln9q
ULUngsXMpbIMoHFg1ePpT2lniiHtQhDC88ZW1J5yeVZN1cfJhRsByyKlkzG3Dcy7
5wKonbVz1LSYpIDq/DNNP0eINzuVNc1rxqRXPGAh6NHnmZN5MqxLNyu0DSX9oJHm
KbbtLQ==
=FQCu
-----END PGP PUBLIC KEY BLOCK-----
\end{lstlisting}

\subsection{Computed key}
\begin{lstlisting}
-----BEGIN PGP PUBLIC KEY BLOCK-----

mQEOBF0zhx4BCAD0iC55GShgArPzKZdn6we2mrkxS8o8fI7HdWXnLoRAta1Q1j9z
U+mYU9tGYb0yjqIARjjUK2qnq0CsGZJpB65AhbrOsDLv6dmI68tuiT4xjmzcM6+6
6M/EY5E5hYpZeurDJ8rSYrfu/v9P0bV78VnM17uUl4PTyEBmbJsE9kszq4MXd5yz
G+ibBPBvnBWSlRQzyCFMgTL8gQH36XJUj7jgxTKoJWzli/QtNcMOjwSQtEZxFWVK
+qsJjYUClpg9Kkow1LqJmEc1Tj98EipypDtdwNQhn4wHQVgu7SliaNNmrR8MpP9o
gL9hhfafx5FsE7v4zHhzu+ubR+c7KnUlwJyfABkBhI5atQAeSm9obiBEb2UgPGNo
YW5nZV9tZUBlbWFpbC5jb20+
=
-----END PGP PUBLIC KEY BLOCK-----
\end{lstlisting}

\newpage

\section{NYSIIS - 1 Static}
\label{appendix:nysiis_1_static}

\subsection{Controlled Key}
\begin{lstlisting}
-----BEGIN PGP PUBLIC KEY BLOCK-----

mI0EXS8o2AEEAKQmxot2Z5XFrNjg4zXBfvLsTzoO3vZeEI9l8+U2X8aq5KVQJpPU
BRSfqQuMZvA1zh3vH3w4I/vR4nXCNw/hPCi0B/DckUe6IJPZJfMiQfT1Xt16ZEms
W53CBiNqnt5xruo4wKwoKGuHELtPcD4K/iBIFt6n3UTi+R8xU4WMc+PjABEBAAG0
HUF1dG9nZW5lcmF0ZWQgS2V5IDx1c2VyQENWRUQ+iNQEEwEKAD4WIQTFXc8VXOqe
13ib81dcbeH3Z8A13wUCXS8o2AIbLwUJAeJb2AULCQgHAgYVCgkICwIEFgIDAQIe
AQIXgAAKCRBcbeH3Z8A13wItA/kBbfc6kxtLumDQiGpcIzgWmM9UhCn0V8ny9iB2
gt4V8bImpC3qicj6+KywsEsvFIgsJO0UcNeoREKanpXxPf6T7tvBUTLQjPJFedsn
J57996JiZ9COQ4lEg2AvfAKy1alKrMs0Xy4hWrHY9cJWs1a7KpbrsioHd4vDRgMS
2yRn/w==
=krrv
-----END PGP PUBLIC KEY BLOCK-----
\end{lstlisting}

\subsection{Computed key}
\begin{lstlisting}
-----BEGIN PGP PUBLIC KEY BLOCK-----

mQEOBF0yA4QBCAC78Lk2k9BdUEufpqll4D60iAuYVrxVs3lpjwP2hpztBFFw8vY4
26uW1C7MPo9l4Oteks5626hwAAgI1/YHFwWpBDdTWYEnVGrjv5v0Juz5IZc7kiwp
Ui2XWJ7I3VpE5nljFyzaQMGaS5Cvmtco5cjin1Tx486jieGBt0loQQdtiJ3gEAFT
XlUvwIQSdYWY2F8xWKg5aJoDFEZN74KfwQqB7hx/53VHyzOJGygnvvpMcJaX3c8E
V3t7n90WvXJWND7bn6HIJrnFKyRJsKHneL2n0wGd/N0NLCuLTvNK+uQ7sXGNZox1
Qa4a1guiSdaq9NO5mX+tOYox3/onzpshJgVDABkBvckntQAeSm9obiBEb2UgPGNo
YW5nZV9tZUBlbWFpbC5jb20+
=
-----END PGP PUBLIC KEY BLOCK-----
\end{lstlisting}

\newpage

\section{NYSIIS - 2 Static}
\label{appendix:nysiis_2_static}

\subsection{Controlled Key}
\begin{lstlisting}
-----BEGIN PGP PUBLIC KEY BLOCK-----

mI0EXS8s+gEEANkph53UuQkNVFhYimrzH0bXBp/Y3InV9p5MP09TZxtTOf44JBPp
MCFMvgFJpVK3D5PQ2h7fW1lNb41THkiGC7Jk+F7gd+beqH5kaaKmC+RNAF0aZ5av
2al5JvtHT0Z8/1r8bxfp33WLmQXK0/9EwL4HvERRGh7wZ5WemvCrvZPXABEBAAG0
HUF1dG9nZW5lcmF0ZWQgS2V5IDx1c2VyQENWRUQ+iNQEEwEKAD4WIQSjUV9vxlLP
LPE78Z2TulybLqhYHAUCXS8s+gIbLwUJAeJXtgULCQgHAgYVCgkICwIEFgIDAQIe
AQIXgAAKCRCTulybLqhYHONrBAC8uh/K2Adq/o/Qe/EVM5JpT6ETE1WlAj2iUMWB
MAmdjgiQlRyBCTd4Lt9xFsbnvQG9pj4GVHosXjVpuA/1NxyS+DE4EkBDzXA3NuR8
0djhyD27VSsn2q1j8bB1wxO7sLbjo30SIBG2wUaft5DVJiBGG69kh4pt+aKKOcjb
LLPldw==
=dCl/
-----END PGP PUBLIC KEY BLOCK-----
\end{lstlisting}

\subsection{Computed key}
\begin{lstlisting}
-----BEGIN PGP PUBLIC KEY BLOCK-----

mQEOBF0wxGEBCACsqQN92G3IizweQHw42OhUM5ZySKTvzPqVhAeCb2ETSBDZFbGN
lHYPa6dytXwLbU6kdlt8kgxbk75iPJsCSPLAUzKvlVd2rkTu5xtKPFqDTRvbjOjf
Ik8zJimXJGZNYVrMUqkfPR2FtVC4VUJXtML8voj/1lcLxTWRcYjuNQuSRNhJQtUP
ViZMTj97IO894lVg+ciZXEZkzHzlriqeLqwTupQxvzqjmObmeH1NCPSfGcTSNIBu
cNG/DrXTXsVpZPUilIRsXTtc2FM9iY49Zjv01KdzsJlV97JVyghYFFCEe1xq7Twp
U9TOzypXD4NdTKivFZUNT4jdG2HoJE4IcZ3NABkBh0qYtB5Kb2huIERvZSA8Y2hh
bmdlX21lQGVtYWlsLmNvbT4=
=0nU6
-----END PGP PUBLIC KEY BLOCK-----
\end{lstlisting}