### 2019-08-13 .archive/Documents/paper_notes/pages/action_needed.tex
Changes; - Archived some documents - Deleted unused files

\section{Action Needed! Helping Users Find and Complete the
Authentication Ceremony in Signal}

Papers overall aim is to study and then improve the ease and provide motivation for users to complete the authentication ceremony in Signal IM.
The authors apply what they refer to as ``Opinionated design'' to naturally assist users in completing the authentication ceremony.

\textit{``Two recent papers demonstrated that
with some instruction about the ceremony itself [8] or the
importance of comparing keys [20], users can successfully
find and use the authentication ceremony. However, users
still took an inordinate amount of time—over 11 minutes on
average—to find and complete the ceremony [20].''}

That paper showed increases in the number of users actually completing the ceremony.

Work was inspired by a study by Vazirpour et al \textit{``Is that you, Alice? a usability study of the authentication ceremony of secure messaging''} [COMPARE RESULTS]. They also borrow some methodology from this work.

The paper's actual modification are inspired by work from \textit{``When SIGNAL hits the fan: On the usability and security of state-of-the-art secure mobile messaging''}