### 2019-08-13 .archive/Documents/paper_notes/pages/druken_bishop.tex

\section{The drunken bishop: An analysis of the OpenSSH
fingerprint visualization algorithm}

The aim of this research is to study the OpenSSH Visual Host Key. The authors claim that the scheme was heuristically designed and has very little backing in terms of formal research, this paper aims to create a foundation for this research.

OpenSSH uses MD5 to generate a 128-bit fingerprint of the Server's key, this is then split into 64 2-bit pairs. These are then traversed byte wise in a right to left over the bit pairs. Each pair defines the direction of travel. If the snake reaches a position multiple times a pre define character is added depending on the number of visits (for example, 6 times == "B")

The paper also discusses ways to create collisions:

In conclusive results and more of an inital PoC

\begin{itemize}
    \item Brute force
    \item Graph theory
    \item Brute forcing a visual set
\end{itemize}