### 2019-08-14 Writeup/Thesis/diagrams/bloom_filter.tex
Changes; - Added frame for conference paper - Moved both projects into a shared folder

\begin{tikzpicture}[
    square/.style={regular polygon,regular polygon sides=4, minimum size=1.4cm}
]

    \node at (0,0) [square,draw] (box_0) {$1$};
    \node at (1,0) [square,draw] (box_1) {$1$};
    \node at (2,0) [square,draw] (box_2) {$1$};
    \node at (3,0) [square,draw] (box_3) {$0$};
    \node at (4,0) [square,draw] (box_4) {$0$};
    \node at (5,0) [square,draw] (box_5) {$0$};
    \node at (6,0) [square,draw] (box_6) {$1$};
    \node at (7,0) [square,draw] (box_7) {$0$};
    \node at (8,0) [square,draw] (box_8) {$1$};
    \node at (9,0) [square,draw] (box_9) {$0$};
    \node at (10,0) [square,draw] (box_10) {$1$};
    \node at (11,0) [square,draw] (box_11) {$0$};

    \node[] (set) [above of=box_5, yshift=2cm] {$\{x, y, z\}$};
    \node[] (set_x) [left of=set, xshift=0.65cm, yshift=-0.15cm] {};
    \node[] (set_z) [right of=set, xshift=-0.65cm, yshift=-0.15cm] {};
    \node[] (new_set) [below of=set, yshift=-4.5cm] {$n$};

    % X
    \draw [->, very thick, blue] (set_x) to [out=-90, in=90, xshift=2cm] (box_6);
    \draw [->, very thick, blue] (set_x) to [out=-90, in=90, xshift=2cm] (box_2);

    % Y
    \draw [->, very thick, red] (set) to [out=-90, in=90] (box_0);
    \draw [->, very thick, red] (set) to [out=-90, in=90] (box_8);

    % Z
    \draw [->, very thick, purple] (set_z) to [out=-90, in=90, xshift=2cm] (box_1);
    \draw [->, very thick, purple] (set_z) to [out=-90, in=90, xshift=2cm] (box_10);

    % N
    \draw [->, thick] (new_set) to [out=90, in=-90, xshift=2cm] (box_1);
    \draw [->, thick] (new_set) to [out=90, in=-90, xshift=2cm] (box_7);
\end{tikzpicture}