### 2019-08-15 Writeup/Thesis/diagrams/graphs/speed.tex
Changes; - More corrections made - Added a speed comparison graph

\begin{tikzpicture}[every axis plot/.append style={thick}]
    \begin{axis}[
        width=\linewidth,
        height=10cm,
        grid=major,
        xmin=1, xmax=40,
        % ymin=0, ymax=100,
        xlabel=Number of Keys,
        ylabel=Hashing speed (MH/s)
        ]
        \addplot table [x index=0, y index=1, mark=none, search path=data, col sep=comma]{scallion.csv};
        \addlegendentry{Scallion}

        \addplot table [x index=0, y index=1, mark=none, search path=data, col sep=comma]{green.csv};
        \addlegendentry{GreenOnion}	
     \end{axis}
\end{tikzpicture}