### 2019-08-20 Writeup/Thesis/code/settings/OpenCL.tex

% ======================================================================== %
% Defines the highlighting, style and environment for the OpenCL language
% ========================================================================%

\lstdefinelanguage{OpenCL}{
  keywords={int, __kernel,kernel,__local,local,__global,global,__constant,constant,__private,private, char2,char3,char4,char8,char16, uchar2,uchar3,uchar4,uchar8,uchar16, short2,short3,short4,short8,short16, ushort2,ushort3,ushort4,ushort8,ushort16, int2,int3,int4,int8,int16, uint2,uint3,uint4,uint8,uint16, long2,long3,long4,long8,long16, ulong2,ulong3,ulong4,ulong8,ulong16, float2,float3,float4,float8,float16, image2d_t,image3d_t,sampler_t,event_t, bool2,bool3,bool4,bool8,bool16, half2,half3,half4,half8,half16, quad,quad2,quad3,quad4,quad8,quad16, complex,imaginary, uint, long, bool, for, if},
  keywordstyle=\color{blue}\bfseries,
  ndkeywords={get_global_id, class, export, boolean, throw, implements, import, this},
  ndkeywordstyle=\color{black}\bfseries,
  comment=[l]{//},
  morecomment=[s]{/*}{*/},
  commentstyle=\itshape\color{OliveGreen}\ttfamily,
  stringstyle=\color{red}\ttfamily,
  morestring=[b]',
  morestring=[b]"
}

\lstdefinestyle{OpenCLStyle}{
    commentstyle=\color{codegreen},
    keywordstyle={\bf \color{blue}},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\footnotesize\ttfamily,
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=left, 
    columns=fixed,                                   
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=7,
    numberstyle=\tiny, 
    numbersep=8pt,
    language=OpenCL,
}

\newtcblisting{OpenCLCode}[2][]{
	top=0cm,
	bottom=0cm,
	left=0.5cm,
	right=0cm,
	colback=white,
	colframe=black,
    arc=1pt, outer arc=1pt,
    listing only, 
    listing style=OpenCLStyle,
}