### 2019-08-26 Writeup/Thesis/code/ExponentIncrement.tex

\begin{adjustbox}{width=\textwidth,height=15cm,keepaspectratio}
\begin{Code}[OpenCLStyle]
///The exponent is split across two words
// i.e
//
// W[3] = XX EE EE EE
// W[4] = EE XX XX XX
//
int newExponent = get_global_id(0) + \
                    ORIGINAL_EXPONENT_VALUE;

// Moves the value to the right i.e. 0x01ffff -> 0x0001ff
// Diving by "0x100" moves the hex value 3 digits
// to the right
int exponentMaskLeft = ((int)(ORIGINAL_EXPONENT_VALUE / 0x100)) ^ ((int)(newExponent / 0x100));

W[3] = W[3] ^ exponentMaskLeft;

// Grabs the last byte of each exponent to apply to 
// the next word
int lastByte = (int)(ORIGINAL_EXPONENT_VALUE / 0x100) *
    0x100 ^ ORIGINAL_EXPONENT_VALUE;

int lastByteNew = (int)(newExponent / 0x100) * 0x100 ^ newExponent;

// Moves it into position as the MSB of a 32-bit integer 
// i.e 0xff -> 0xff000000
int exponenetMaskRight = (lastByte * 0x1000000) ^ 
    (lastByteNew * 0x1000000);

W[4] = W[4] ^ exponenetMaskRight;
\end{Code}
\end{adjustbox}