### 2019-08-28 Writeup/Thesis/chapters/abstract.tex

Many encrypted connections require the comparison of a fingerprint to protect against eavesdropping. A substantial amount of past work has aimed to propose fingerprint representations that work better with human limitations. "Trustwords" proposed by pEp is an example of such a scheme, where the fingerprint is encoded as words in an attempt to improve usability.

This work implements an attack on Trustwords and quantifies its effectiveness on more than 400 participants.  A tool called GreenOnion was designed to assist in quantifying attack feasibility. GreenOnion improved substantially on a similar tool's ability to search for matches concurrently. Our findings show a substantial increase in attack success compared to related literature. We believe this increase is due to leveraging the design flaws in the Trustword scheme.