### 2019-08-29 Writeup/Thesis/graphs/speed.tex
Changes; - All fixes implemented from supervisor feedback

\begin{tikzpicture}[every axis plot/.append style={thick}]
    \begin{axis}[
        width=\linewidth,
        height=10cm,
        grid=major,
        xmin=1, xmax=2513,
        % ymin=0, ymax=100,
        x tick label style={rotate=45},
        x label style={at={(axis description cs:0.5,-0.1)},anchor=north},
        xlabel=Number of Keys,
        ylabel=Hashing speed (MH/s)
        ]
        \addplot[color=blue] table [x index=0, y index=1, mark=none, search path=data, col sep=comma]{scallion.csv};
        \addlegendentry{Scallion}

        \addplot[color=red] table [x index=0, y index=1, mark=none, search path=data, col sep=comma]{fscallion.csv};
        \addlegendentry{Scallion (Improved)}

        \addplot[color=green] table [x index=0, y index=1, mark=none, search path=data, col sep=comma]{green.csv};
        \addlegendentry{GreenOnion}	
     \end{axis}
\end{tikzpicture}