### 2019-08-30 Writeup/Thesis/code/GoogleAppScript.tex

\begin{adjustbox}{width=\textwidth,keepaspectratio}
  \begin{Code}[JavascriptStyle]
vvar ALGOS = ["Soundex", "Metaphone", ...];
var ALGO_SIZES = [763777, 412916, ...];

var QUESTION_PER_ALGO = 5;

function updateForm(ss, form) {

  // Obtains a list of the items present on the form
  formItems = form.getItems();

  var algo_index = 0;
  var scale_index = 0;
  for (var i = 0; i < formItems.length; i++)
  {
    if (formItems[i].getType() == "SCALE")
    {
      var rnd_index = Math.floor(Math.random() * 
      ALGO_SIZES[algo_index]) + 1;

      // Obtains the Google Sheets object
      var sheetname = ALGOS[algo_index];
      var sheet = ss.getSheetByName(sheetname);
      var range = sheet.getRange(rnd_index, 1)
      var values = range.getValues();

      // Re-samples from the connected Google Sheets
      _updateScaleTitle(formItems[i], values[0]);

      scale_index++;

      // Moves on to the next metric when a group size 
      // has been completed
      if (scale_index != 0)
      {
        if (scale_index % QUESTION_PER_ALGO == 0) {
            algo_index++;
        }
      }
    }
  }
}
  \end{Code}
\end{adjustbox}