### 2019-08-30 Writeup/Thesis/chapters/abstract.tex
Changes; - TTS run over the entire paper and corrections added

Many encrypted connections require the comparison of a fingerprint to protect against eavesdropping. A substantial amount of past work has aimed to propose fingerprint representations that work better with human limitations. "Trustwords" proposed by pEp is an example of such a scheme, where the fingerprint is encoded as words in an attempt to improve usability.

This works main aim is to asses if Trustword's recommend minimum number of four words is sufficient.
In order to achieve this goal, this work implements an attack on Trustwords and quantifies its effectiveness on more than 400 participants.  A tool called GreenOnion was designed to assist in quantifying attack feasibility. GreenOnion improved substantially on a similar tool's ability to search for matches concurrently where it was able to search for over 1.5 million concurrent keys. Our findings show a substantial increase in attack success compared to related literature. We believe this increase is due to levering the design flaws in the Trustword scheme. Therefore, we believe a minimum of four Trustwords is insufficient to provide even a basic level of security.